---
title: "Cervical_cancer_violetgo"
author: "Xinyu Wang"
date: "2023-05-03"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1 Data preprocessing:
### Packages
```{r}
library(ggplot2)
library(dplyr)
library(Boruta)
library(DataExplorer)
library(rmarkdown)
library(flexdashboard)
library(readr)
library(GGally)
```
### Load Data
```{r}
cervical <- read.csv("risk_factors_cervical_cancer.csv")
head(cervical,5)# first 5 rows of data
```
### Columns
```{r}
colnames(cervical) #names of variables
```

### data info
```{r}
glimpse(cervical)
summary(cervical)
```

### delete characters like "?" since this can lead to char
```{r}
for(i in 1:ncol(cervical)) { 
  for (j in 1:nrow(cervical)){
  if (cervical[j,i] == "?"){
  cervical[j , i] <- NA}
  else next
  }}
head(cervical)
```
### to numeric
```{r}
# ro numeric
cervical <- apply(cervical,2,as.numeric)
cervical <- round(cervical,digits = 2)
cervical <- as.data.frame(cervical)
# some cols to factors
cervical$Smokes <- as.factor(cervical$Smokes)
cervical$Hormonal.Contraceptives <- as.factor(cervical$Hormonal.Contraceptives)
cervical$IUD <- as.factor(cervical$IUD)

cervical$STDs <- as.factor(cervical$STDs)
cervical$STDs.condylomatosis <- as.factor(cervical$STDs.condylomatosis)
cervical$STDs.vaginal.condylomatosis <- as.factor(cervical$STDs.vaginal.condylomatosis)
cervical$STDs.vulvo.perineal.condylomatosis <- as.factor(cervical$STDs.condylomatosis)
cervical$STDs.condylomatosis <- as.factor(cervical$STDs.vulvo.perineal.condylomatosis)
cervical$STDs.syphilis  <- as.factor(cervical$STDs.syphilis)
cervical$STDs.pelvic.inflammatory.disease <- as.factor(cervical$STDs.pelvic.inflammatory.disease)
cervical$STDs.genital.herpes <- as.factor(cervical$STDs.genital.herpes)
cervical$STDs.molluscum.contagiosum <- as.factor(cervical$STDs.molluscum.contagiosum)
cervical$STDs.AIDS <- as.factor(cervical$STDs.AIDS)
cervical$STDs.HIV <- as.factor(cervical$STDs.HIV)
cervical$STDs.Hepatitis.B <- as.factor(cervical$STDs.Hepatitis.B)
cervical$STDs.HPV <- as.factor(cervical$STDs.HPV)
# Dx changed as bool
cervical$Dx.Cancer <- as.factor(cervical$Dx.Cancer)
cervical$Dx.CIN <- as.factor(cervical$Dx.CIN)
cervical$Dx.HPV <- as.factor(cervical$Dx)
cervical$Hinselmann <- as.factor(cervical$Hinselmann)
cervical$Schiller <- as.factor(cervical$Schiller)
cervical$Citology <- as.factor(cervical$Citology)
cervical$Biopsy <- as.factor(cervical$Biopsy)
head(cervical,5)
```
### missing values
```{r}
# delete if too much missing values and leave it as it is if not
na_prop <- colMeans(is.na(cervical))
cervical <- cervical[,na_prop <= 0.8]
cervical_storage <- cervical
cervical_withNA <- cervical
cervical_withMean <- cervical
```

```{r}
#filling the value NA
for (col_name in names(cervical_withMean)) {
  # if it
  if (is.numeric(cervical_withMean[[col_name]])) {
    mean_value <- mean(cervical_withMean[[col_name]], na.rm = TRUE)
    cervical_withMean[[col_name]][is.na(cervical_withMean[[col_name]])] <- mean_value
  }
  # if col is factor, then use random sampling
  else if (is.factor(cervical_withMean[[col_name]])) {
    set.seed(42) 
    cervical_withMean[[col_name]][is.na(cervical_withMean[[col_name]])] <- sample(cervical_withMean[[col_name]][!is.na(cervical_withMean[[col_name]])], 
                                                     sum(is.na(cervical_withMean[[col_name]])), 
                                                     replace = TRUE)
  }
}  
```

```{r}
summary(cervical_withMean)
```

### standardization
```{r}
# Remove columns with all values equal to zero before standardization
cervical_without_zero_cols <- cervical_withMean[, !colSums(cervical_withMean == 0, na.rm = TRUE) == nrow(cervical_withMean)]

# Standardize the numeric columns
for (col_name in names(cervical_without_zero_cols)) {
  if (is.numeric(cervical_without_zero_cols[[col_name]])) {
    min_value <- min(cervical_without_zero_cols[[col_name]], na.rm = TRUE)
    max_value <- max(cervical_without_zero_cols[[col_name]], na.rm = TRUE)
    cervical_without_zero_cols[[col_name]] <- (cervical_without_zero_cols[[col_name]] - min_value) / (max_value - min_value)
  }
}

# Update the dataset
cervical_std <- cervical_without_zero_cols

# Display the first few rows of the standardized dataset
head(cervical_std)

```

## 2 EDA 
```{r}
library(ggtext)

prop_NA <- function(x) { mean(is.na(x)) }
missdata <- sapply(cervical, prop_NA)
missdata <- data.frame(variable = names(missdata), prop_NA = missdata)
missdata <- missdata[order(missdata$prop_NA, decreasing = TRUE), ]

ggplot(missdata, aes(x = variable, y = prop_NA, fill = variable)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9)) + 
  scale_fill_viridis_d(guide = FALSE) +
  labs(x = "Variable", y = "Proportion of Missing Values", 
       title = "Proportion of Missing Values for Each Variable in Cervical Cancer Risk Dataset") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
plot_str(cervical)
plot_bar(cervical)
plot_qq(cervical)

```

## 3 Feature Selection

### Define the target response
```{r}
Cervical <- cervical_std
Dx_dt <- Cervical[,c("Hinselmann","Schiller","Citology","Biopsy")]
Cervical$Outcome <- apply(Dx_dt, 1, FUN = function(x) {return(sum(as.numeric(x)))})
```

### Proportion of the Outcome with not 0 and 0
```{r}
# Convert the "Outcome" column to numeric data type
Cervical$Outcome <- as.numeric(Cervical$Outcome)
Cervical$Outcome <- as.numeric(Cervical$Outcome)

prop <- nrow(Cervical[Cervical$Outcome==0,])/nrow(Cervical[Cervical$Outcome>0,])
cat("The ratio between control and case is", prop, "for the whole data suppose case when outcome is larger than 0.\n")
```
### Split into training and testing set
```{r}
Npop <- nrow(Cervical)
test_ind <- sample(Npop, Npop/5)
cervical_te <- Cervical[test_ind,]
cervical_tr <- Cervical[-test_ind,]
```

### Downsampling for training set and test set(feature selection)
## Downsampling for traing set
```{r}
# Load the necessary library
library(dplyr)

# Separate cases and controls
cases <- cervical_tr[cervical_tr$Outcome != 0,]
controls <- cervical_tr[cervical_tr$Outcome == 0,]

# Function to downsample controls for different ratios
downsample_controls <- function(ratio) {
  n_controls <- nrow(cases) * ratio
  return(sample_n(controls, n_controls))
}

# Perform downsampling
controls_1_2 <- downsample_controls(2)
controls_1_3 <- downsample_controls(3)
controls_1_4 <- downsample_controls(4)
controls_1_5 <- downsample_controls(5)

# Combine cases and downsampled controls for each ratio
cervical_tr_1_2 <- rbind(cases, controls_1_2)
cervical_tr_1_3 <- rbind(cases, controls_1_3)
cervical_tr_1_4 <- rbind(cases, controls_1_4)
cervical_tr_1_5 <- rbind(cases, controls_1_5)

# Shuffle the rows (if needed)
cervical_tr_1_2 <- cervical_tr_1_2[sample(nrow(cervical_tr_1_2)),]
cervical_tr_1_3 <- cervical_tr_1_3[sample(nrow(cervical_tr_1_3)),]
cervical_tr_1_4 <- cervical_tr_1_4[sample(nrow(cervical_tr_1_4)),]
cervical_tr_1_5 <- cervical_tr_1_5[sample(nrow(cervical_tr_1_5)),]
#We choose cervical_tr_1_3 for model(feature) selection.
```

##Downsampling for test set
```{r}
# Separate cases and controls
cases_te <- cervical_te[cervical_te$Outcome != 0,]
controls_te <- cervical_te[cervical_te$Outcome == 0,]

# Function to downsample controls for different ratios
downsample_controls <- function(ratio) {
  n_controls <- nrow(cases) * ratio
  return(sample_n(controls, n_controls))
}

# Perform downsampling
controls_1_2_te <- downsample_controls(2)
controls_1_3_te <- downsample_controls(3)
controls_1_4_te <- downsample_controls(4)
controls_1_5_te <- downsample_controls(5)

# Combine cases and downsampled controls for each ratio
cervical_te_1_2 <- rbind(cases, controls_1_2)
cervical_te_1_3 <- rbind(cases, controls_1_3)
cervical_te_1_4 <- rbind(cases, controls_1_4)
cervical_te_1_5 <- rbind(cases, controls_1_5)

#  Shuffle the rows (if needed)
cervical_te_1_2 <- cervical_te_1_2[sample(nrow(cervical_tr_1_2)),]
cervical_te_1_3 <- cervical_te_1_3[sample(nrow(cervical_tr_1_3)),]
cervical_te_1_4 <- cervical_te_1_4[sample(nrow(cervical_tr_1_4)),]
cervical_te_1_5 <- cervical_te_1_5[sample(nrow(cervical_tr_1_5)),]
#We choose cervical_tr_1_3 for model(feature) selection.
```
### Delete the columns with all zeros in cervical_tr_1_3
```{r}
# Remove columns with all zeros
cervical_tr_1_2 <- cervical_tr_1_2[, colSums(cervical_tr_1_2 != 0) > 0]
cervical_tr_1_3 <- cervical_tr_1_3[, colSums(cervical_tr_1_3 != 0) > 0]
cervical_tr_1_4 <- cervical_tr_1_4[, colSums(cervical_tr_1_4 != 0) > 0]
cervical_tr_1_5 <- cervical_tr_1_5[, colSums(cervical_tr_1_5 != 0) > 0]
```
### LASSO feature selection
```{r}
X_tr <- cervical_tr_1_4[,1:24]
Y_tr <- cervical_tr_1_4[,"Outcome"]
X_te <- cervical_te_1_4[,1:24]
Y_te <- cervical_te_1_4[,"Outcome"]
X_class <- data.frame(sapply(X_tr, class))
```

```{r}
library(glmnet)
# input X should be a matrix
numeric_df <- X_tr[, sapply(X_tr, is.numeric)]
numeric_mat <- as.matrix(numeric_df)
factor_df <- sapply(X_tr[, sapply(X_tr, is.factor)], as.numeric)-1
factor_mat <- as.matrix(factor_df)
Xmat_tr <- cbind(numeric_mat, factor_mat)
```


```{r}
numeric_df <- X_te[, sapply(X_te, is.numeric)]
numeric_mat <- as.matrix(numeric_df)
factor_df <- sapply(X_te[, sapply(X_te, is.factor)], as.numeric)-1
factor_mat <- as.matrix(factor_df)
Xmat_te <- cbind(numeric_mat, factor_mat)
```


```{r}
lasso_fit <- glmnet(Xmat_tr, as.numeric(Y_tr), family = "gaussian", alpha = 1)

```


```{r}
cv.out <- cv.glmnet(Xmat_tr, as.numeric(Y_tr), family = "gaussian")
plot(cv.out)
bestlam <- cv.out$lambda.min
lasso_pred <- predict(lasso_fit , s = bestlam , newx = Xmat_te, type = "response")
yhat <- ifelse(lasso_pred >1, 1, 0)
table(yhat)
conf.mat <- table(yhat, Y_te)
accuracy <- sum(diag(conf.mat))/sum(conf.mat)
accuracy
print(bestlam)# best lambda roughly 0.03, 80 accuracy
```

#### Extract non-zero variables
```{r}
all_coef <- coef(lasso_fit, s = bestlam)
nonzero_coef <- all_coef[all_coef[, 1] != 0, ]
cat("Non-zero coefficients:\n")
nonzero_coef
```


## 4 Model training

### linear regression

```{r}
# Function to fit models for a given dataset
linear_1_2 <- lm(Outcome ~ Num.of.pregnancies + Smokes..years. + Smokes..packs.year. +Hormonal.Contraceptives..years. + IUD + STDs.condylomatosis +STDs.vaginal.condylomatosis + STDs.vulvo.perineal.condylomatosis +
STDs.syphilis + STDs.genital.herpes + STDs.HIV + STDs.Hepatitis.B, data = cervical_tr_1_2)
linear_1_3 <- lm(Outcome ~ Num.of.pregnancies + Smokes..years. + Smokes..packs.year. +Hormonal.Contraceptives..years. + IUD + STDs.condylomatosis +STDs.vaginal.condylomatosis + STDs.vulvo.perineal.condylomatosis +
STDs.syphilis + STDs.genital.herpes + STDs.HIV + STDs.Hepatitis.B, data = cervical_tr_1_3)
linear_1_4 <- lm(Outcome ~ Num.of.pregnancies + Smokes..years. + Smokes..packs.year. +Hormonal.Contraceptives..years. + IUD + STDs.condylomatosis +STDs.vaginal.condylomatosis + STDs.vulvo.perineal.condylomatosis +
STDs.syphilis + STDs.genital.herpes + STDs.HIV + STDs.Hepatitis.B, data = cervical_tr_1_4)
linear_1_5 <- lm(Outcome ~ Num.of.pregnancies + Smokes..years. + Smokes..packs.year. +Hormonal.Contraceptives..years. + IUD + STDs.condylomatosis +STDs.vaginal.condylomatosis + STDs.vulvo.perineal.condylomatosis +
STDs.syphilis + STDs.genital.herpes + STDs.HIV + STDs.Hepatitis.B, data = cervical_tr_1_5)

```

### predict
```{r}
linear_pred_1_2 <- predict(linear_1_2, cervical_te_1_2)
linear_pred_1_3 <- predict(linear_1_3, cervical_te_1_3)
linear_pred_1_4 <- predict(linear_1_4, cervical_te_1_4)
linear_pred_1_5 <- predict(linear_1_5, cervical_te_1_5)
```
### MSE&MAE
```{r}
# Calculate MSE for each dataset
mse_1_2 <- mean((cervical_te_1_2$Outcome - linear_pred_1_2)^2)
mse_1_3 <- mean((cervical_te_1_3$Outcome - linear_pred_1_3)^2)
mse_1_4 <- mean((cervical_te_1_4$Outcome - linear_pred_1_4)^2)
mse_1_5 <- mean((cervical_te_1_5$Outcome - linear_pred_1_5)^2)
#MAE
mae_1_2 <- mean(abs(cervical_te_1_2$Outcome - linear_pred_1_2))
mae_1_3 <- mean(abs(cervical_te_1_3$Outcome - linear_pred_1_3))
mae_1_4 <- mean(abs(cervical_te_1_4$Outcome - linear_pred_1_4))
mae_1_5 <- mean(abs(cervical_te_1_5$Outcome - linear_pred_1_5))
```
### plot MSE for each 
```{r}
# Create a vector with MSE values
mse_values <- c(mse_1_2, mse_1_3, mse_1_4, mse_1_5)

# Create a vector with dataset names
datasets <- c("1:2", "1:3", "1:4", "1:5")

# Create a barplot for MSE values
barplot(mse_values, names.arg = datasets, xlab = "Datasets", ylab = "MSE", main = "MSE for Each Dataset", col = "skyblue")
```
### Do the same for random forest model

```{r}
# Fit random forest models
library(randomForest)
rf_1_2 <- randomForest(Outcome ~ Num.of.pregnancies + Smokes..years. + Smokes..packs.year. +Hormonal.Contraceptives..years. + IUD + STDs.condylomatosis +STDs.vaginal.condylomatosis + STDs.vulvo.perineal.condylomatosis +
STDs.syphilis + STDs.genital.herpes + STDs.HIV + STDs.Hepatitis.B, data = cervical_tr_1_2,importance=TRUE)
rf_1_3 <- randomForest(Outcome ~ Num.of.pregnancies + Smokes..years. + Smokes..packs.year. +Hormonal.Contraceptives..years. + IUD + STDs.condylomatosis +STDs.vaginal.condylomatosis + STDs.vulvo.perineal.condylomatosis +
STDs.syphilis + STDs.genital.herpes + STDs.HIV + STDs.Hepatitis.B, data = cervical_tr_1_3,importance=TRUE)
rf_1_4 <- randomForest(Outcome ~ Num.of.pregnancies + Smokes..years. + Smokes..packs.year. +Hormonal.Contraceptives..years. + IUD + STDs.condylomatosis +STDs.vaginal.condylomatosis + STDs.vulvo.perineal.condylomatosis +
STDs.syphilis + STDs.genital.herpes + STDs.HIV + STDs.Hepatitis.B, data = cervical_tr_1_4,importance=TRUE)
rf_1_5 <- randomForest(Outcome ~ Num.of.pregnancies + Smokes..years. + Smokes..packs.year. +Hormonal.Contraceptives..years. + IUD + STDs.condylomatosis +STDs.vaginal.condylomatosis + STDs.vulvo.perineal.condylomatosis +
STDs.syphilis + STDs.genital.herpes + STDs.HIV + STDs.Hepatitis.B, data = cervical_tr_1_5,importance=TRUE)
```


```{r}
# Make predictions
rf_pred_1_2 <- predict(rf_1_2, cervical_te_1_2)
rf_pred_1_3 <- predict(rf_1_3, cervical_te_1_3)
rf_pred_1_4 <- predict(rf_1_4, cervical_te_1_4)
rf_pred_1_5 <- predict(rf_1_5, cervical_te_1_5)

# Calculate MSE
mse_rf_1_2 <- mean((cervical_te_1_2$Outcome - rf_pred_1_2)^2)
mse_rf_1_3 <- mean((cervical_te_1_3$Outcome - rf_pred_1_3)^2)
mse_rf_1_4 <- mean((cervical_te_1_4$Outcome - rf_pred_1_4)^2)
mse_rf_1_5 <- mean((cervical_te_1_5$Outcome - rf_pred_1_5)^2)

#MAE
mae_rf_1_2 <- mean(abs(cervical_te_1_2$Outcome - rf_pred_1_2))
mae_rf_1_3 <- mean(abs(cervical_te_1_2$Outcome - rf_pred_1_3))
mae_rf_1_4 <- mean(abs(cervical_te_1_2$Outcome - rf_pred_1_4))
mae_rf_1_5 <- mean(abs(cervical_te_1_2$Outcome - rf_pred_1_5))

# Create a vector with MSE values
mse_rf_values <- c(mse_rf_1_2, mse_rf_1_3, mse_rf_1_4, mse_rf_1_5)


importance(rf_1_2)
varImpPlot(rf_1_2)

importance(rf_1_3)
varImpPlot(rf_1_3)

importance(rf_1_4)
varImpPlot(rf_1_4)

importance(rf_1_5)
varImpPlot(rf_1_5)

# Create a bar plot for MSE values
barplot(mse_rf_values, names.arg = datasets, xlab = "Datasets", ylab = "MSE", main = "Random Forest MSE for Each Dataset", col = "skyblue")

```

### Do the same for the SVM
```{r}
# Load the necessary library
library(e1071)

# Fit SVM models
svm_1_2 <- svm(Outcome ~ Num.of.pregnancies + Smokes..years. + Smokes..packs.year. +Hormonal.Contraceptives..years. + IUD + STDs.condylomatosis +STDs.vaginal.condylomatosis + STDs.vulvo.perineal.condylomatosis +
STDs.syphilis + STDs.genital.herpes + STDs.HIV + STDs.Hepatitis.B, data = cervical_tr_1_2)
svm_1_3 <- svm(Outcome ~ Num.of.pregnancies + Smokes..years. + Smokes..packs.year. +Hormonal.Contraceptives..years. + IUD + STDs.condylomatosis +STDs.vaginal.condylomatosis + STDs.vulvo.perineal.condylomatosis +
STDs.syphilis + STDs.genital.herpes + STDs.HIV + STDs.Hepatitis.B, data = cervical_tr_1_3)
svm_1_4 <- svm(Outcome ~ Num.of.pregnancies + Smokes..years. + Smokes..packs.year. +Hormonal.Contraceptives..years. + IUD + STDs.condylomatosis +STDs.vaginal.condylomatosis + STDs.vulvo.perineal.condylomatosis +
STDs.syphilis + STDs.genital.herpes + STDs.HIV + STDs.Hepatitis.B, data = cervical_tr_1_4)
svm_1_5 <- svm(Outcome ~ Num.of.pregnancies + Smokes..years. + Smokes..packs.year. +Hormonal.Contraceptives..years. + IUD + STDs.condylomatosis +STDs.vaginal.condylomatosis + STDs.vulvo.perineal.condylomatosis +
STDs.syphilis + STDs.genital.herpes + STDs.HIV + STDs.Hepatitis.B, data = cervical_tr_1_5)

# Make predictions
svm_pred_1_2 <- predict(svm_1_2, cervical_te_1_2)
svm_pred_1_3 <- predict(svm_1_3, cervical_te_1_3)
svm_pred_1_4 <- predict(svm_1_4, cervical_te_1_4)
svm_pred_1_5 <- predict(svm_1_5, cervical_te_1_5)

# Calculate MSE
mse_svm_1_2 <- mean((cervical_te_1_2$Outcome - svm_pred_1_2)^2)
mse_svm_1_3 <- mean((cervical_te_1_3$Outcome - svm_pred_1_3)^2)
mse_svm_1_4 <- mean((cervical_te_1_4$Outcome - svm_pred_1_4)^2)
mse_svm_1_5 <- mean((cervical_te_1_5$Outcome - svm_pred_1_5)^2)

#MAE
mae_svm_1_2 <- mean(abs(cervical_te_1_2$Outcome - svm_pred_1_2))
mae_svm_1_3 <- mean(abs(cervical_te_1_2$Outcome - svm_pred_1_3))
mae_svm_1_4 <- mean(abs(cervical_te_1_2$Outcome - svm_pred_1_4))
mae_svm_1_5 <- mean(abs(cervical_te_1_2$Outcome - svm_pred_1_5))

# Create a vector with MSE values
mse_svm_values <- c(mse_svm_1_2, mse_svm_1_3, mse_svm_1_4, mse_svm_1_5)

# Create a bar plot for MSE values
barplot(mse_svm_values, names.arg = datasets, xlab = "Datasets", ylab = "MSE", main = "SVM MSE for Each Dataset", col = "skyblue")

```
### gbm
```{r}
# Load the necessary libraries
library(gbm)
library(caret)

# Fit GBM models
gbm_1_2 <- gbm(Outcome ~ Num.of.pregnancies + Smokes..years. + Smokes..packs.year. +Hormonal.Contraceptives..years. + IUD + STDs.condylomatosis +STDs.vaginal.condylomatosis + STDs.vulvo.perineal.condylomatosis +
STDs.syphilis + STDs.genital.herpes + STDs.HIV + STDs.Hepatitis.B, data = cervical_tr_1_2, distribution = "gaussian", n.trees = 5000, interaction.depth = 1, cv.folds = 5)
gbm_1_3 <- gbm(Outcome ~ Num.of.pregnancies + Smokes..years. + Smokes..packs.year. +Hormonal.Contraceptives..years. + IUD + STDs.condylomatosis +STDs.vaginal.condylomatosis + STDs.vulvo.perineal.condylomatosis +
STDs.syphilis + STDs.genital.herpes + STDs.HIV + STDs.Hepatitis.B, data = cervical_tr_1_3, distribution = "gaussian", n.trees = 5000, interaction.depth = 1, cv.folds = 5)
gbm_1_4 <- gbm(Outcome ~ Num.of.pregnancies + Smokes..years. + Smokes..packs.year. +Hormonal.Contraceptives..years. + IUD + STDs.condylomatosis +STDs.vaginal.condylomatosis + STDs.vulvo.perineal.condylomatosis +
STDs.syphilis + STDs.genital.herpes + STDs.HIV + STDs.Hepatitis.B, data = cervical_tr_1_4, distribution = "gaussian", n.trees = 5000, interaction.depth = 1, cv.folds = 5)
gbm_1_5 <- gbm(Outcome ~ Num.of.pregnancies + Smokes..years. + Smokes..packs.year. +Hormonal.Contraceptives..years. + IUD + STDs.condylomatosis +STDs.vaginal.condylomatosis + STDs.vulvo.perineal.condylomatosis +
STDs.syphilis + STDs.genital.herpes + STDs.HIV + STDs.Hepatitis.B, data = cervical_tr_1_5, distribution = "gaussian", n.trees = 5000, interaction.depth = 1, cv.folds = 5)

best_n_tress_1_2 <- which.min(gbm_1_2$cv.error)
summary(gbm_1_2$cv.error)

best_n_tress_1_3 <- which.min(gbm_1_3$cv.error)
summary(gbm_1_3$cv.error)

best_n_tress_1_4 <- which.min(gbm_1_4$cv.error)
summary(gbm_1_4$cv.error)

best_n_tress_1_5 <- which.min(gbm_1_5$cv.error)
summary(gbm_1_5$cv.error)

```

```{r}
# Make predictions
gbm_pred_1_2 <- predict(gbm_1_2, cervical_te_1_2, n.trees =best_n_tress_1_2)
gbm_pred_1_3 <- predict(gbm_1_3, cervical_te_1_3, n.trees =best_n_tress_1_3)
gbm_pred_1_4 <- predict(gbm_1_4, cervical_te_1_4, n.trees =best_n_tress_1_4)
gbm_pred_1_5 <- predict(gbm_1_5, cervical_te_1_5, n.trees =best_n_tress_1_5)

# Calculate MSE
mse_gbm_1_2 <- mean((cervical_te_1_2$Outcome - gbm_pred_1_2)^2)
mse_gbm_1_3 <- mean((cervical_te_1_3$Outcome - gbm_pred_1_3)^2)
mse_gbm_1_4 <- mean((cervical_te_1_4$Outcome - gbm_pred_1_4)^2)
mse_gbm_1_5 <- mean((cervical_te_1_5$Outcome - gbm_pred_1_5)^2)

# MAE
mae_gbm_1_2 <- mean(abs(cervical_te_1_2$Outcome - gbm_pred_1_2))
mae_gbm_1_3 <- mean(abs(cervical_te_1_2$Outcome - gbm_pred_1_3))
mae_gbm_1_4 <- mean(abs(cervical_te_1_2$Outcome - gbm_pred_1_4))
mae_gbm_1_5 <- mean(abs(cervical_te_1_2$Outcome - gbm_pred_1_5))

# Create a vector with MSE values
mse_gbm_values <- c(mse_gbm_1_2, mse_gbm_1_3, mse_gbm_1_4, mse_gbm_1_5)

# Create a bar plot for MSE values
barplot(mse_gbm_values, names.arg = datasets, xlab = "Datasets", ylab = "MSE", main = "GBM MSE for Each Dataset", col = "skyblue")

```
### KNN
```{r}
library(caret)

# Specify the range of k values
k_range <- 1:20

# Initialize a list to store MSE values for each dataset
mse_values_list <- list()

# Iterate over the datasets
datasets1 <- list(cervical_tr_1_2, cervical_tr_1_3, cervical_tr_1_4, cervical_tr_1_5)
test_datasets <- list(cervical_te_1_2, cervical_te_1_3, cervical_te_1_4, cervical_te_1_5)

for (i in 1:4) {
  # Initialize a vector to store MSE values for each k
  mse_values <- numeric(length(k_range))
  
  # Iterate over the range of k values
  for (k in k_range) {
    # Fit the k-NN regression model
    knnreg_model <- knnreg(Outcome ~ Num.of.pregnancies + Smokes..years. + Smokes..packs.year. +Hormonal.Contraceptives..years. + IUD + STDs.condylomatosis +STDs.vaginal.condylomatosis + STDs.vulvo.perineal.condylomatosis +
STDs.syphilis + STDs.genital.herpes + STDs.HIV + STDs.Hepatitis.B, data = datasets1[[i]], k)
    
    # Make predictions on the testing dataset
    knnreg_pred <- predict(knnreg_model, test_datasets[[i]])
    
    # Calculate MSE for the current k value
    mse_values[k] <- mean((test_datasets[[i]]$Outcome - knnreg_pred)^2)
  }
  mse_values_list[[i]] <- mse_values
}

# Find the optimal k value with the minimum MSE for each dataset
optimal_k_values <- sapply(mse_values_list, which.min)

# Visualize MSE values for each k and dataset
par(mfrow = c(2, 2))
for (i in 1:length(datasets1)) {
  plot(k_range, mse_values_list[[i]], type = "b", xlab = "k Values", ylab = "MSE", main = paste("MSE for Different k Values in k-NN Regression (Dataset", i, ")"), col = "blue")
  abline(v = optimal_k_values[i], col = "red", lty = 2)
}

```

```{r}
df1 <- data.frame(
  
  k_range = k_range,
  mse_values = mse_values_list
)

# Print the data frame
print(df1)
```

```{r}
# Load the necessary libraries
library(caret)

# Fit knnreg models
knnreg_model_1_2 <- knnreg(Outcome ~ Num.of.pregnancies + Smokes..years. + Smokes..packs.year. +Hormonal.Contraceptives..years. + IUD + STDs.condylomatosis +STDs.vaginal.condylomatosis + STDs.vulvo.perineal.condylomatosis +
STDs.syphilis + STDs.genital.herpes + STDs.HIV + STDs.Hepatitis.B, data = cervical_tr_1_2,k=2)
knnreg_model_1_3 <- knnreg(Outcome ~ Num.of.pregnancies + Smokes..years. + Smokes..packs.year. +Hormonal.Contraceptives..years. + IUD + STDs.condylomatosis +STDs.vaginal.condylomatosis + STDs.vulvo.perineal.condylomatosis +
STDs.syphilis + STDs.genital.herpes + STDs.HIV + STDs.Hepatitis.B, data = cervical_tr_1_3,k=8)
knnreg_model_1_4 <- knnreg(Outcome ~ Num.of.pregnancies + Smokes..years. + Smokes..packs.year. +Hormonal.Contraceptives..years. + IUD + STDs.condylomatosis +STDs.vaginal.condylomatosis + STDs.vulvo.perineal.condylomatosis +
STDs.syphilis + STDs.genital.herpes + STDs.HIV + STDs.Hepatitis.B, data = cervical_tr_1_4,k=5)
knnreg_model_1_5 <- knnreg(Outcome ~ Num.of.pregnancies + Smokes..years. + Smokes..packs.year. +Hormonal.Contraceptives..years. + IUD + STDs.condylomatosis +STDs.vaginal.condylomatosis + STDs.vulvo.perineal.condylomatosis +
STDs.syphilis + STDs.genital.herpes + STDs.HIV + STDs.Hepatitis.B, data = cervical_tr_1_5,k=5)

# Make predictions
knnreg_pred_1_2 <- predict(knnreg_model_1_2, cervical_te_1_2)
knnreg_pred_1_3 <- predict(knnreg_model_1_3, cervical_te_1_3)
knnreg_pred_1_4 <- predict(knnreg_model_1_4, cervical_te_1_4)
knnreg_pred_1_5 <- predict(knnreg_model_1_5, cervical_te_1_5)

# Calculate MSE
mse_knnreg_1_2 <- mean((cervical_te_1_2$Outcome - knnreg_pred_1_2)^2)
mse_knnreg_1_3 <- mean((cervical_te_1_3$Outcome - knnreg_pred_1_3)^2)
mse_knnreg_1_4 <- mean((cervical_te_1_4$Outcome - knnreg_pred_1_4)^2)
mse_knnreg_1_5 <- mean((cervical_te_1_5$Outcome - knnreg_pred_1_5)^2)

# MAE
mae_gbm_1_2 <- mean(abs(cervical_te_1_2$Outcome - gbm_pred_1_2))
mae_gbm_1_3 <- mean(abs(cervical_te_1_2$Outcome - gbm_pred_1_3))
mae_gbm_1_4 <- mean(abs(cervical_te_1_2$Outcome - gbm_pred_1_4))
mae_gbm_1_5 <- mean(abs(cervical_te_1_2$Outcome - gbm_pred_1_5))

# Create a vector with MSE values
mse_knnreg_values <- c(mse_knnreg_1_2, mse_knnreg_1_3, mse_knnreg_1_4, mse_knnreg_1_5)

# Create a bar plot for MSE values
barplot(mse_knnreg_values, names.arg = datasets, xlab = "Datasets", ylab = "MSE", main = "KNN Regression MSE for Each Dataset", col = "skyblue")

```
### MSE 
```{r}
# Create a line plot for MSE values of Linear Regression, Random Forest, and SVM
# Prepare the data
mse <- function(true, pred) {
  return(mean((true - pred)^2))
}
mse_values <- matrix(nrow = 5, ncol = 4)
mse_values[1,] <- c(mse(cervical_te_1_2$Outcome, linear_pred_1_2),
                    mse(cervical_te_1_3$Outcome, linear_pred_1_3),
                    mse(cervical_te_1_4$Outcome, linear_pred_1_4),
                    mse(cervical_te_1_5$Outcome, linear_pred_1_5))

mse_values[2,] <- c(mse(cervical_te_1_2$Outcome, rf_pred_1_2),
                    mse(cervical_te_1_3$Outcome, rf_pred_1_3),
                    mse(cervical_te_1_4$Outcome, rf_pred_1_4),
                    mse(cervical_te_1_5$Outcome, rf_pred_1_5))
mse_values[3,] <- c(mse(cervical_te_1_2$Outcome, knnreg_pred_1_2),
                    mse(cervical_te_1_3$Outcome, knnreg_pred_1_3),
                    mse(cervical_te_1_4$Outcome, knnreg_pred_1_4),
                    mse(cervical_te_1_5$Outcome, knnreg_pred_1_5))
mse_values[4,] <- c(mse(cervical_te_1_2$Outcome, svm_pred_1_2),
                    mse(cervical_te_1_3$Outcome, svm_pred_1_3),
                    mse(cervical_te_1_4$Outcome, svm_pred_1_4),
                    mse(cervical_te_1_5$Outcome, svm_pred_1_5))
mse_values[5,] <- c(mse(cervical_te_1_2$Outcome, gbm_pred_1_2),
                    mse(cervical_te_1_3$Outcome, gbm_pred_1_3),
                    mse(cervical_te_1_4$Outcome, gbm_pred_1_4),
                    mse(cervical_te_1_5$Outcome, gbm_pred_1_5))

datasets <- c("1:2", "1:3", "1:4", "1:5")
```

```{r}
library(grid)
# Create a line plot for MSE values of Linear Regression, Random Forest, and SVM
plot(mse_values[1,], type = "o", xaxt = "n", xlab = "Datasets", ylab = "MSE", main = "MSE for Each Dataset", col = "blue", pch = 19, ylim = range(mse_values),lwd=5)
lines(mse_values[2,], type = "o", col = "red", pch = 19,lwd=5)
lines(mse_values[3,], type = "o", col = "darkgreen", pch = 19,lwd=5)
lines(mse_values[4,], type = "o", col = "purple", pch = 19,lwd=5)
lines(mse_values[5,], type = "o", col = "orange", pch = 19,lwd=5)

# Add axis labels
axis(1, at = 1:length(datasets), labels = datasets)

# Add gridlines
grid(nx = NULL, ny = NULL, col = "gray", lty = "dotted", lwd = par("lwd"), equilogs = TRUE)

# Add legend
legend("topright", legend = c("Linear Regression", "Random Forest", "knn", "svm", "Gradient Boost Machine"), col = c("blue", "red", "darkgreen", "purple", "orange"), pch = 19, lty = 1, bty = "n")

```
### RMSE
```{r}
# Calculate RMSE values for each model and dataset
rmse_values <- sqrt(mse_values)

# Create a line plot for RMSE values of Linear Regression, Random Forest, and SVM
plot(rmse_values[1,], type = "o", xaxt = "n", xlab = "Datasets", ylab = "RMSE", main = "RMSE for Each Dataset", col = "blue", pch = 19, ylim = range(mse_values),lwd=5)
lines(rmse_values[2,], type = "o", col = "red", pch = 19,lwd=5)
lines(rmse_values[3,], type = "o", col = "darkgreen", pch = 19,lwd=5)
lines(rmse_values[4,], type = "o", col = "purple", pch = 19,lwd=5)
lines(rmse_values[5,], type = "o", col = "orange", pch = 19,lwd=5)

# Add axis labels
axis(1, at = 1:length(datasets), labels = datasets)

# Add gridlines
grid(nx = NULL, ny = NULL, col = "gray", lty = "dotted", lwd = par("lwd"), equilogs = TRUE)

# Add legend
legend("topright", legend = c("Linear Regression", "Random Forest", "knn", "svm", "Gradient Boost Machine"), col = c("blue", "red", "darkgreen", "purple", "orange"),pch = 19, lty = 1, bty = "n")


```
### MAE
```{r}
# Create a line plot for MSE values of Linear Regression, Random Forest, and SVM
# Prepare the data
mae <- function(true, pred) {
  return(mean(abs(true - pred)))
}
mae_values <- matrix(nrow = 5, ncol = 4)
mae_values[1,] <- c(mae(cervical_te_1_2$Outcome, linear_pred_1_2),
                    mae(cervical_te_1_3$Outcome, linear_pred_1_3),
                    mae(cervical_te_1_4$Outcome, linear_pred_1_4),
                    mae(cervical_te_1_5$Outcome, linear_pred_1_5))

mae_values[2,] <- c(mae(cervical_te_1_2$Outcome, rf_pred_1_2),
                    mae(cervical_te_1_3$Outcome, rf_pred_1_3),
                    mae(cervical_te_1_4$Outcome, rf_pred_1_4),
                    mae(cervical_te_1_5$Outcome, rf_pred_1_5))
mae_values[3,] <- c(mae(cervical_te_1_2$Outcome, knnreg_pred_1_2),
                    mae(cervical_te_1_3$Outcome, knnreg_pred_1_3),
                    mae(cervical_te_1_4$Outcome, knnreg_pred_1_4),
                    mae(cervical_te_1_5$Outcome, knnreg_pred_1_5))
mae_values[4,] <- c(mae(cervical_te_1_2$Outcome, svm_pred_1_2),
                    mae(cervical_te_1_3$Outcome, svm_pred_1_3),
                    mae(cervical_te_1_4$Outcome, svm_pred_1_4),
                    mae(cervical_te_1_5$Outcome, svm_pred_1_5))
mae_values[5,] <- c(mae(cervical_te_1_2$Outcome, gbm_pred_1_2),
                    mae(cervical_te_1_3$Outcome, gbm_pred_1_3),
                    mae(cervical_te_1_4$Outcome, gbm_pred_1_4),
                    mae(cervical_te_1_5$Outcome, gbm_pred_1_5))

datasets <- c("1:2", "1:3", "1:4", "1:5")
```

```{r}
library(grid)
# Create a line plot for MSE values of Linear Regression, Random Forest, and SVM
plot(mae_values[1,], type = "o", xaxt = "n", xlab = "Datasets", ylab = "MAE", main = "MAE for Each Dataset", col = "blue", pch = 19, ylim = range(mae_values),lwd=5)
lines(mae_values[2,], type = "o", col = "red", pch = 19,lwd=5)
lines(mae_values[3,], type = "o", col = "darkgreen", pch = 19,lwd=5)
lines(mae_values[4,], type = "o", col = "purple", pch = 19,lwd=5)
lines(mae_values[5,], type = "o", col = "orange", pch = 19,lwd=5)

# Add axis labels
axis(1, at = 1:length(datasets), labels = datasets)

# Add gridlines
grid(nx = NULL, ny = NULL, col = "gray", lty = "dotted", lwd = par("lwd"), equilogs = TRUE)

# Add legend
legend("topright", legend = c("Linear Regression", "Random Forest", "knn", "svm", "Gradient Boost Machine"), col = c("blue", "red", "darkgreen", "purple", "orange"), pch = 19, lty = 1, bty = "n")

```
### table: MSE,RMSE,MAE
```{r}
# Create a data frame containing the accuracy values
df <- data.frame(
  Model = rep(c("Linear Regression", "Random Forest", "knn","svm","Gradient Boost Machine"), each = 4),
  Dataset = rep(datasets, 5),
  MSE = c(mse_values),
  RMSE = c(rmse_values),
  MAE = c(mae_values)
)

# Round the accuracy values to 3 decimal places
df$MSE <- round(df$MSE, 3)
df$RMSE <- round(df$RMSE, 3)
df$MAE <- round(df$MAE, 3)
# Print the data frame
print(df)

```
### CV(coefficient of variation) to measure stability of downsampling
```{r}
# Add CV columns for each metric
df$CV_MSE <- NA
df$CV_RMSE <- NA
df$CV_MAE <- NA

# Calculate CV for each model
unique_models <- unique(df$Model)
for (model in unique_models) {
  model_indices <- df$Model == model
  df$CV_MSE[model_indices] <- sd(df$MSE[model_indices]) / mean(df$MSE[model_indices])
  df$CV_RMSE[model_indices] <- sd(df$RMSE[model_indices]) / mean(df$RMSE[model_indices])
  df$CV_MAE[model_indices] <- sd(df$MAE[model_indices]) / mean(df$MAE[model_indices])
}

# Round the CV values to 3 decimal places
df$CV_MSE <- round(df$CV_MSE, 3)
df$CV_RMSE <- round(df$CV_RMSE, 3)
df$CV_MAE <- round(df$CV_MAE, 3)

# Print the data frame
print(df)

```

```{r}
# Calculate accuracy
accuracy <- function(true, pred) {
  return(mean(true == round(pred))) # true value equals to the rounded predicted value
}

acc_values <- matrix(nrow = 5, ncol = 4)
acc_values[1,] <- c(accuracy(cervical_te_1_2$Outcome, linear_pred_1_2),
                    accuracy(cervical_te_1_3$Outcome, linear_pred_1_3),
                    accuracy(cervical_te_1_4$Outcome, linear_pred_1_4),
                    accuracy(cervical_te_1_5$Outcome, linear_pred_1_5))

acc_values[2,] <- c(accuracy(cervical_te_1_2$Outcome, rf_pred_1_2),
                    accuracy(cervical_te_1_3$Outcome, rf_pred_1_3),
                    accuracy(cervical_te_1_4$Outcome, rf_pred_1_4),
                    accuracy(cervical_te_1_5$Outcome, rf_pred_1_5))

acc_values[3,] <- c(accuracy(cervical_te_1_2$Outcome, knnreg_pred_1_2),
                    accuracy(cervical_te_1_3$Outcome, knnreg_pred_1_3),
                    accuracy(cervical_te_1_4$Outcome, knnreg_pred_1_4),
                    accuracy(cervical_te_1_5$Outcome, knnreg_pred_1_5))

acc_values[4,] <- c(accuracy(cervical_te_1_2$Outcome, svm_pred_1_2),
                    accuracy(cervical_te_1_3$Outcome, svm_pred_1_3),
                    accuracy(cervical_te_1_4$Outcome, svm_pred_1_4),
                    accuracy(cervical_te_1_5$Outcome, svm_pred_1_5))

acc_values[5,] <- c(accuracy(cervical_te_1_2$Outcome, gbm_pred_1_2),
                    accuracy(cervical_te_1_3$Outcome, gbm_pred_1_3),
                    accuracy(cervical_te_1_4$Outcome, gbm_pred_1_4),
                    accuracy(cervical_te_1_5$Outcome, gbm_pred_1_5))


# Create a grouped bar plot for accuracy values
barplot(acc_values, beside = TRUE, names.arg = datasets, xlab = "Datasets", ylab = "Accuracy", main = "Accuracy of Each Model", col = c("blue", "red", "green","purple","yellow"), legend.text = c("Linear Regression", "Random Forest", "knn","svm","Gradient Boost Machine"), args.legend = list(x = "topleft", bty = "n"))


```
### table: accuracy
```{r}
# Create a data frame containing the accuracy values
accuracy_df <- data.frame(
  Model = rep(c("Linear Regression", "Random Forest", "knn","svm","Gradient Boost Machine"), each = 4),
  Dataset = rep(datasets, 5),
  Accuracy = c(acc_values)
)

# Round the accuracy values to 3 decimal places
accuracy_df$Accuracy <- round(accuracy_df$Accuracy, 3)

# Print the data frame
print(accuracy_df)
```
